Policing for Profit and the Mobilization of Black Voters

### 2021-03-22 markdown/draft_paper.tex
write

\title{Policing for Profit and the Mobilization of Black Voters}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Fines, Fees, and Turnout}
\author{Jonathan Ben-Menachem\footnote{PhD Student, Columbia University, Department of Sociology (\href{mailto:jb4487@columbia.edu}{\nolinkurl{jb4487@columbia.edu}})} \and Kevin Morris\footnote{PhD Student, CUNY Graduate Center, Department of Sociology (\href{mailto:kmorris@gradcenter.cuny.edu}{\nolinkurl{kmorris@gradcenter.cuny.edu}})}}
\date{March 17, 2021}

\begin{document}
\maketitle
\begin{abstract}
The American criminal legal system is an important site of political socialization: sociologists show how the alienating effect of criminalization manifests as systems avoidance, and political scientists have measured the effects of policing and incarceration on political participation. Despite this burgeoning literature, no research has directly investigated how ticketing for low-level offenses affects political participation. Using data from the Census of Governments and the national voter file, we test the causal effect of fines and fees practices on American voting behavior at the city level. Additionally, we exploit individual-level ticketing data from Hillsborough County, Florida between 2003 and 2019 to determine whether getting a ticket impacts political participation. Nationally, we estimate that each 10 percent increase in city fines and fees collections between 2012 and 2017 is associated with a roughly 0.6 percentage point decrease in nonBlack turnout but was not associated with Black turnout. In Tampa, Florida, we estimate being ticketed increases a nonBlack individual's likelihood of voting by 3.7 percentage points, and that the effect is larger for Black voters.

@@

It wasn't just a Ferguson problem, or even a Missouri problem. American cities' reliance on fines and fees revenue increased significantly following the 2008 recession --- as local tax revenues dropped and tax increases became less politically viable, jurisdictions increased the amounts of fines and fees and imposed them more frequently in order to fund government services (\protect\hyperlink{ref-Singla2020}{Singla, Kirschner, and Stone 2020}; \protect\hyperlink{ref-Harris2017}{Harris et al. 2017}).

Work exploring how criminalization directly and indirectly influences political participation has exploded in recent years. This work, however, has largely focused on the effects of highly consequential contact with the criminal legal system such as incarceration and felony convictions (\protect\hyperlink{ref-Burch2014}{Burch 2014}; \protect\hyperlink{ref-Lee2014}{Lee, Porter, and Comfort 2014}). Although some work has leveraged survey data and qualitative approaches to interrogate lower-level police interactions --- such as ticketing --- our project represents the first use of individual-level administrative data to identify the causal effect of ticketing on voter behavior. We begin by joining fees and fines data from municipalities around the country in 2012 and 2017 with a novel dataset estimating racial turnout for each municipality in the country. While we find that increased fees and fines likely cause lower turnout among non-Black Americans, we find no effect of fees and fines on Black turnout.

After exploring national data, we use individual-level traffic stop data from Hillsborough County, Florida, to identify the turnout patterns of voters who were ticketed between the 2012 and 2018 elections. By matching individuals who were ticketed to other, similar voters who were not ticketed and running a difference-in-differences model, we estimate the causal effect of these tickets on turnout. Using self-reported race from the Florida registered voter file, we identify large, \emph{positive} treatment effects for all voters --- effects that were especially large for Black voters..

We demonstrate that police ticketing practices --- the most widespread form of police contact in America --- influence the political behavior of Americans, with distinct effects for Black and non-Black voters. This finding supports and extends Hannah Walker's theory of criminal legal contact is mobilizing / less demobilizing for nonwhite political participation (\protect\hyperlink{ref-Walker2020}{Walker 2020}, \protect\hyperlink{ref-Walker2014}{2014}). On an individual level, we also find that voters who were ticketed turn out at higher rates in subsequent elections, and that this effect was especially large for Black individuals. This finding complicates Vesla Weaver and Amy Lerman's theory of low-level criminal legal contact discouraging political participation (\protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Our findings are relevant for interdisciplinary scholars of crime, race, politics, political economy, social movements, and policing.

@@

Tthe role of fines and fees in the context of felony disenfranchisement has recently garnered national attentiontaken on greater political significance. After Floridians passed Amendment 4 voted to restore voting rights to individuals with felony convictions in 2018, Florida Republicans passed legislation requiring these individuals to pay off the fines and fees associated with their original sentence. Because Florida does not maintain good records on who owes how much, this has proved to be a major hurdle for potential voters. Despite a dissenting judge arguing that ````{[}h{]}ad Florida wanted to create a system to obstruct, impede, and impair the ability of felons to vote under Amendment 4, it could not have come up with a better one,'' an en bank ruling by the U.S. Court of Appeals for the 11th Circuit upheld this legislation.\footnote{Jones et al.~v. DeSantis et al., 4:19cv300-RH/MJF (United States Court of Appeals for the Eleventh Circuit).}

@@

Using existing national survey data and extended interviews, Vesla Weaver and Amy Lerman found that trust in government and willingness to vote decrease as individuals progress through increasingly intense levels of criminal legal contact (questioned by police, arrested, convicted, incarcerated) (\protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Weaver and Lerman also investigate the potential role of criminal legal contact in facilitating a collective political consciousness among Black Americans, or the concept of a ``linked fate'' where group members believe that group-level phenomena affect all group members (\protect\hyperlink{ref-Cox2019}{Cox 2019}) of all Black Americans. Their interview data lead them to conclude that ``a custodial linked fate is not the building block of greater activism and a mobilized racial identity; instead, it is demobilizing'' --- and interview respondents who had been incarcerated reported that withdrawal from public institutions and democratic participation constituted a form of self-preservation (\protect\hyperlink{ref-Lerman2014}{Lerman and Weaver 2014}).

@@

It may be the case that Black and non-Black Americans employ different self-preservation strategies, however. Ariel White's analysis of how short jail sentences affect voting in Harris County, Texas found that Black defendant turnout in subsequent elections dropped by 13 percentage points --- but the effect on white turnout was not statistically significant (\protect\hyperlink{ref-White2019}{White 2019}).

Proximal exposure to criminalization --- ``having a loved one who is a custodial citizen without yourself having had contact'' (\protect\hyperlink{ref-Walker2017}{Walker and García-Castañon 2017}) --- also influences Americans' perceptions of government and political behavior, and the emerging literature has produced mixed findings of its effect on civic engagement. Using survey and ethnographic interview data, \protect\hyperlink{ref-Lee2014}{Lee, Porter, and Comfort} (\protect\hyperlink{ref-Lee2014}{2014}) found that adult children of incarcerated parents are less likely to vote, report lower trust in government, and are less likely to engage in other prosocial behaviors like community service. \protect\hyperlink{ref-Sugie2015}{Sugie} (\protect\hyperlink{ref-Sugie2015}{2015}) used survey data to show that partners of formerly incarcerated men are less likely to vote, but not less likely to participate in religious or civic institutions.

Hannah Walker's research specifically tests differences between the effects of proximal and personal criminal legal contact on political behavior, and her findings underscore how a ``linked fate'' political analysis could catalyze distinct political behavior among Black voters (\protect\hyperlink{ref-Walker2014}{Walker 2014}). Using data from a 2012 survey of Washington state residents, she found that as the degree of an individual's proximity to a criminalized person increases, so does their likelihood of participating in political activities (such as voting, signing a petition, volunteering, donating to politicians, etc.). Importantly, Walker found that nonwhite people who had experienced \emph{either} proximal or personal contact \emph{at any level of intensity}, including being stopped and questioned by police, were politically mobilized --- and the same held true for whites, but with a much smaller effect size. Walker theorizes that this finding suggests that nonwhite people view criminalization ``as a form of systemic political threat,'' an analysis comparable to ``linked fate.'' Notably, Walker's survey questions excluded ``minor traffic stops,'' the primary focus of this paper.

Ariel White leveraged administrative court records from Harris County, Texas, to identify voters who lived within individuals who were charged with a misdemeanor offense and arrested, convicted, or jailed (\protect\hyperlink{ref-White2019}{White 2019}). She found that living with someone who was charged but not convicted in the 30 day period preceding an election increased voter turnout by 3 percentage points, whereas ``convicted but not jailed'' decreased turnout by 6 percentage points and ``convicted and jailed'' decreased turnout by 12 percentage points. This provides further support for the theory that increasing intensity of criminalization could change the direction of the effect of criminalization on political behavior --- and, in fact, that lower-level contact can mobilize even as higher-level contact demobilizes.

Although geographic or spatial analyses of how criminalization affects political socialization technically move beyond proximal contact as an extant conceptual framework, we argue that geographic proximity should work similarly to social network proximity --- if one neighborhood is heavily policed, people who live in that neighborhood are more likely to have social ties who have experienced policing. Traci Burch found that North Carolina block groups with higher concentrations of people who had experienced incarceration or criminal legal supervision saw substantial reductions in voter turnout, and confirmed her findings at the individual level (\protect\hyperlink{ref-Burch2014}{Burch 2014}). \protect\hyperlink{ref-Morris2020}{Morris} (\protect\hyperlink{ref-Morris2020}{2020}) similarly finds that neighborhoods with more active voters lost to felony disenfranchisement turned out at lower rates in the 2017 mayoral race in NYC, and that this effect was particularly pronounced in Black neighborhoods.

@@

Over the past few years, scientific understanding of the political effects of more serious interactions with the criminal legal system has grown by leaps and bounds. The following table summarizes our understanding of the literature.

Type of criminal legal contact
Personal contact effect on political behavior (individual level)
Proximal contact effect on political behavior (social network/geography)
Stopped
Weaver and Lerman find chilling effect; Walker finds mobilizing effect
Laniyonu and Walker find mobilizing effect
Arrested
Weaver and Lerman find chilling effect; Walker finds mobilizing effect
Walker finds mobilizing effect
Convicted / Supervision
Burch, Weaver and Lerman find chilling effect; Walker finds mobilizing effect
Burch, Lee, Sugie, White, and Morris find chilling effect; Walker finds mobilizing effect
Incarcerated
Burch, White, Weaver and Lerman find chilling effect; Walker finds mobilizing effect
Burch, Lee, Sugie, White, and Morris find chilling effect; Walker finds mobilizing effect

\caption{\label{tab:lit-table}\label{tab:lit-table} Summary of Literature}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}}
\toprule
Type of criminal legal contact & Personal contact effect on political behavior (individual level) & Proximal contact effect on political behavior (social network/geography)\\
\midrule
\cellcolor{gray!6}{Stopped} & \cellcolor{gray!6}{Weaver and Lerman find chilling effect; Walker finds mobilizing effect} & \cellcolor{gray!6}{Laniyonu and Walker find mobilizing effect}\\
Arrested & Weaver and Lerman find chilling effect; Walker finds mobilizing effect & Walker finds mobilizing effect\\
\cellcolor{gray!6}{Convicted / Supervision} & \cellcolor{gray!6}{Burch, Weaver and Lerman find chilling effect; Walker finds mobilizing effect} & \cellcolor{gray!6}{Burch, Lee, Sugie, White, and Morris find chilling effect; Walker finds mobilizing effect}\\
Incarcerated & Burch, White, Weaver and Lerman find chilling effect; Walker finds mobilizing effect & Burch, Lee, Sugie, White, and Morris find chilling effect; Walker finds mobilizing effect\\
\bottomrule
\end{tabular}
\end{table}
\end{singlespace}

Most of the aforementioned studies were conducted using data collected before the rise of the Black Lives Matter movement in 2013, which could be construed as a significant development for the political socialization of Black Americans with respect to policing and incarceration. The movement garnered national attention and momentum during the Ferguson protests in 2014, where racially discriminatory ticketing practices were particularly egregious. Though racist police violence was and is a focal point for many Black Lives Matter actions, local chapters frequently use policing as an entry point into a wider set of demands regarding the material security of Black Americans and their marginalization in American politics --- and these demands affect police practices. One recent study found that census places that saw Black Lives Matter protests experienced a 15 to 20 percent decrease in police killings over the following five year period (\protect\hyperlink{ref-Campbell2021}{Campbell 2021}). It's possible that any potentially mobilizing effects of criminal legal contact became stronger in the 2010s --- in other words, Black Lives Matter could have facilitated a more robust and prevalent linked-fate political analysis among Black Americans during the time frame of our analysis. That said, \protect\hyperlink{ref-Laniyonu2019}{Laniyonu} (\protect\hyperlink{ref-Laniyonu2019}{2019}) did demonstrate mobilization in Black neighborhoods before the beginning of the Black Lives Matter movement.

@@

In the 20th century, residential segregation simultaneously fueled the racial wealth gap and ensured that criminalization was both spatially and racially concentrated (\protect\hyperlink{ref-Rothstein2017}{Rothstein 2017}; \protect\hyperlink{ref-Hinton2021}{Hinton and Cook 2021}). At the level of federal policy, racial threat and whites' attribution of criminality to Blackness (\protect\hyperlink{ref-Muhammad2010}{Muhammad 2010}) prompted a shift ``from the War on Poverty to the War on Crime'' wherein Black communities' demands for material security and safety were selectively amplified to promote increased and harsher policing and incarceration (\protect\hyperlink{ref-Hinton2017}{Hinton 2017}; \protect\hyperlink{ref-Vargas2017a}{Vargas and McHarris 2017}). Ruth Wilson Gilmore explains how mid to late 20th-century whites ``demanded that the state become leaner and meaner, except when directed to do otherwise'' --- in other words, Keynesian welfare guarantees were undermined by the racist myth of ``welfare queens'' and supplanted by expanded police power (\protect\hyperlink{ref-Gilmore2007}{Gilmore 2007, 53}).

The 2008 financial crisis solidified the connection between fiscal austerity and policing, prompting a decline in tax revenues that devastated state and local governments --- but law enforcement agencies were largely spared from cuts. At the local level, this means that police budgets today constitute about thirty percent of city budgets nationwide, and cities where residents have lower median incomes dedicate a larger proportion of spending to policing (\protect\hyperlink{ref-Friedman2020}{Friedman 2020}). In municipalities with substantial Black populations, then, people who are most heavily affected by criminalization are expected to foot the bill for their own mistreatment.

@@

In 2020, a Federal Reserve survey found that five percent of white respondents reported ``unpaid legal expenses, fines, or court costs,'' whereas twelve percent of Black respondents reported such debt (\protect\hyperlink{ref-BoardofGovernorsoftheFederalReserveSystem2020}{Governors of the Federal Reserve System 2020}). Across the country, city government reliance on fines and fees revenue is associated with larger Black populations (\protect\hyperlink{ref-UnitedStatesCommissiononCivilRights2017}{United States Commission on Civil Rights 2017}). Although major urban centers collect high dollar amounts of fines and fees, some of the most egregious ticketing practices measured on a per-capita basis are concentrated in suburbs or smaller cities with municipal courts --- Ferguson being a prominent example (\protect\hyperlink{ref-UnitedStatesDepartmentofJusticeCivilRightsDivision2015}{United States Department of Justice Civil Rights Division 2015}; \protect\hyperlink{ref-Mughan2020}{Mughan 2020}). The positive association between cities' reliance on fines and fees and the proportion of Black residents is significantly diminished when Black communities are represented by at least one Black legislator, however (\protect\hyperlink{ref-Sances2017}{Sances and You 2017}).

@@

Research regarding police stops and searches in response to local fiscal crisis shows how decreasing local revenues affect racist police practices. One study of all U.S. city governments found that local fiscal stress was related to increased ``revenue-generating arrests'' (e.g.~arrests involving civil asset forfeiture) and that Black and Hispanic people were arrested at higher rates (\protect\hyperlink{ref-Makowsky2009}{Makowsky and Stratmann 2009}). A study of four million North Carolina traffic stops found that fiscal stress caused a decrease in the probability of white drivers being searched, but no decrease for Black drivers (\protect\hyperlink{ref-Shoub2020}{\textbf{Shoub2020?}}). This suggests that higher reliance on fines and fees results in government officials and police targeting individuals who hold less political power --- whether those individuals are out-of-state motorists or members of local marginalized communities (\protect\hyperlink{ref-Makowsky2009}{Makowsky and Stratmann 2009}).

@@

We merge these geocoded voter files with ``place'' and ``county subdivision'' shapefiles published by the US Census Bureau. By aggregating these individual-level records to the municipal level, we are able to estimate the number of registered voters and the number of ballots cast in every city in the country. By using the racial estimates provided by L2, we can estimate the number of ballots cast by voters of each race in each city. Although each voters' race is modeled based on L2's algorithms --- and is not self-reported --- it is possibly incorrect for some voters. We assume this does not present a problem when records are aggregated to the municipal level.

@@

\emph{Hypothesis 2: The effect of the city-level relationship between fines and voter turnout is less demobilizing or more mobilizing for Black voters, given the historical and explicit racial injustice narratives used to explain these practices..}

@@

To test the effect of ticketing on individual-level turnout, we match the first and last names and dates of birth against the Hillsborough County registered voter file. Here, rather than rely on a snapshot from L2, we use the raw voter file directly from the state. Using a single post-treatment snapshot of the voter file can result in conditioning on a post-treatment status (see \protect\hyperlink{ref-Nyhan2017}{Nyhan, Skovron, and Titiunik 2017}). Instead, we collect snapshots of the voter file following each even-year general election between 2012 and 2018. We thus observe virtually all individuals who were registered to vote at any time during our period of study. Unique voter identification numbers allow us to avoid double-counting voters who are registered in multiple snapshots. We each voter's earliest record, and geocode voters to their home census block groups.

@@

The use of administrative data marks an important step forward in our understanding of how low-level contact with the criminal legal system structures participation. Past work looking at the individual-level effects of low-level contact has relied entirely on survey or qualitative data (e.g. \protect\hyperlink{ref-Walker2014}{Walker 2014}; \protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Their in-depth research allows for the testing of specific psychological mechanisms and personal interpretations of criminal legal contact, but does not allow us to generalize more broadly. As Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010, 821}) note, it may also introduce measurement bias. Our analysis investigates actual voting behavior following actual ticketing events, not reported voting behavior or reported exposure to police ticketing. The administrative data therefore allow us both to sidestep reporting error and to observe the behavior of the 232,568 individuals ticketed over the 6 year period --- a far larger pool than even the most robust surveys.

@@

Figure \ref{fig:did2} makes this effect even clearer. Here, we limit the analysis to voters who were stopped within the 90 days prior to an election (and their controls). We expected that the ticket would be more salient for individuals stopped shortly before the election. This appears to be the case --- \emph{especially for Black voters.}

@@

\begin{singlespace}
\input{"../temp/dind_reg.tex"}
\end{singlespace}

>>>>

\title{Policing for Profit and the Mobilization of Black Voters}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Fines, Fees, and Turnout}
\author{Jonathan Ben-Menachem\footnote{PhD Student, Columbia University, Department of Sociology (\href{mailto:jb4487@columbia.edu}{\nolinkurl{jb4487@columbia.edu}})} \and Kevin Morris\footnote{PhD Student, CUNY Graduate Center, Department of Sociology (\href{mailto:kmorris@gradcenter.cuny.edu}{\nolinkurl{kmorris@gradcenter.cuny.edu}})}}
\date{March 19, 2021}

\begin{document}
\maketitle
\begin{abstract}
The American criminal legal system is an important site of political socialization: sociologists show how the alienating effect of criminalization manifests as systems avoidance, and political scientists have measured the effects of policing and incarceration on political participation. Despite this burgeoning literature, no research has directly investigated how ticketing for low-level offenses affects political participation. Using data from the Census of Governments and the national voter file, we test the causal effect of fines and fees practices on American voting behavior at the city level. Additionally, we exploit individual-level ticketing data from Hillsborough County, Florida between 2003 and 2019 to determine whether getting a ticket impacts political participation. Nationally, we estimate that each 10 percent increase in city fines and fees collections between 2012 and 2017 is associated with a roughly 0.6 percentage point decrease in non-Black turnout but was not associated with Black turnout. In Tampa, Florida, we estimate that being ticketed increases a non-Black individual's likelihood of voting by 3.7 percentage points, and that the effect is larger for Black voters.

@@

It wasn't just a Ferguson problem, or even a Missouri problem. American cities' reliance on fines and fees revenue increased significantly following the 2008 recession---as local tax revenues dropped and tax increases became less politically viable, jurisdictions increased the amounts of fines and fees and imposed them more frequently in order to fund government services (\protect\hyperlink{ref-Singla2020}{Singla, Kirschner, and Stone 2020}; \protect\hyperlink{ref-Harris2017}{Harris et al. 2017}).

Work exploring how criminalization directly and indirectly influences political participation has exploded in recent years. This work, however, has largely focused on the effects of highly consequential contact with the criminal legal system such as incarceration and felony convictions (\protect\hyperlink{ref-Burch2014}{Burch 2014}; \protect\hyperlink{ref-Lee2014}{Lee, Porter, and Comfort 2014}). Although some work has leveraged survey data and qualitative approaches to interrogate lower-level police interactions---such as ticketing---our project represents the first use of individual-level administrative data to identify the causal effect of ticketing on voter behavior. We begin by joining fees and fines data from municipalities around the country in 2012 and 2017 with a novel dataset estimating racial turnout for each municipality in the country. While we find that increased fees and fines likely cause lower turnout among non-Black Americans, we find no effect of fees and fines on Black turnout.

After exploring national data, we use individual-level traffic stop data from Hillsborough County, Florida, to identify the turnout patterns of voters who were ticketed between the 2012 and 2018 elections. By matching individuals who were ticketed to other, similar voters who were not ticketed and running a difference-in-differences model, we estimate the causal effect of these tickets on turnout. Using self-reported race from the Florida registered voter file, we identify large, \emph{positive} treatment effects for all voters---effects that were especially large for Black voters..

We demonstrate that police ticketing practices---the most widespread form of police contact in America---influence the political behavior of Americans, with distinct effects for Black and non-Black voters. This finding supports and extends Hannah Walker's theory of criminal legal contact is mobilizing / less demobilizing for nonwhite political participation (\protect\hyperlink{ref-Walker2020}{Walker 2020}, \protect\hyperlink{ref-Walker2014}{2014}). Our results also complicate Vesla Weaver and Amy Lerman's theory of low-level criminal legal contact discouraging political participation (\protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Our findings are relevant for interdisciplinary scholars of crime, race, politics, political economy, social movements, and policing.

@@

The role of fines and fees in the context of felony disenfranchisement has recently garnered national attention. After Floridians passed Amendment 4 to restore voting rights to individuals with felony convictions in 2018, Florida Republicans passed legislation requiring these individuals to pay off the fines and fees associated with their original sentence. Because Florida does not maintain good records on who owes how much, this has proved to be a major hurdle for potential voters. Despite a dissenting judge arguing that ````{[}h{]}ad Florida wanted to create a system to obstruct, impede, and impair the ability of felons to vote under Amendment 4, it could not have come up with a better one,'' an en banc ruling by the U.S. Court of Appeals for the 11th Circuit upheld this legislation.\footnote{Jones et al.~v. DeSantis et al., 4:19cv300-RH/MJF (United States Court of Appeals for the Eleventh Circuit).}

@@

Using existing national survey data and extended interviews, Vesla Weaver and Amy Lerman found that trust in government and willingness to vote decrease as individuals progress through increasingly intense levels of criminal legal contact (questioned by police, arrested, convicted, incarcerated) (\protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Weaver and Lerman also investigate the potential role of criminal legal contact in facilitating a collective political consciousness among Black Americans, or the concept of a ``linked fate'' where ``one's individual opportunities are linked to the entire racial group'' (\protect\hyperlink{ref-Lerman2014}{Lerman and Weaver 2014, 267}). Their interview data lead them to conclude that while criminal legal contact does increase perceptions of linked fate among Black respondents, ``a custodial linked fate is not the building block of greater activism and a mobilized racial identity; instead, it is demobilizing''---and interview respondents who had been incarcerated reported that withdrawal from public institutions and democratic participation constituted a form of self-preservation (\protect\hyperlink{ref-Lerman2014}{Lerman and Weaver 2014}).

@@

It may be the case that Black and non-Black Americans employ different self-preservation strategies, however. Ariel White's analysis of how short jail sentences affect voting in Harris County, Texas found that Black defendant turnout in subsequent elections dropped by 13 percentage points---but the effect on white turnout was not statistically significant (\protect\hyperlink{ref-White2019}{White 2019a}).

Proximal exposure to criminalization---``having a loved one who is a custodial citizen without yourself having had contact'' (\protect\hyperlink{ref-Walker2017}{Walker and García-Castañon 2017})---also influences Americans' perceptions of government and political behavior, and the emerging literature has produced mixed findings of its effect on civic engagement. Using survey and ethnographic interview data, \protect\hyperlink{ref-Lee2014}{Lee, Porter, and Comfort} (\protect\hyperlink{ref-Lee2014}{2014}) found that adult children of incarcerated parents are less likely to vote, report lower trust in government, and are less likely to engage in other prosocial behaviors like community service. \protect\hyperlink{ref-Sugie2015}{Sugie} (\protect\hyperlink{ref-Sugie2015}{2015}) used survey data to show that partners of formerly incarcerated men are less likely to vote, but not less likely to participate in religious or civic institutions.

Hannah Walker's research specifically tests differences between the effects of proximal and personal criminal legal contact on political behavior, and her findings underscore how a ``linked fate'' political analysis could catalyze distinct political behavior among Black voters (\protect\hyperlink{ref-Walker2014}{Walker 2014}). Using data from a 2012 survey of Washington state residents, she found that as the degree of an individual's proximity to a criminalized person increases, so does their likelihood of participating in political activities other than voting (such as signing a petition, volunteering, donating to politicians, etc.). With respect to voting, she found that personal contact was demobilizing, while proximal contact had no statistically significant effect.

Importantly, Walker found that nonwhite people who had experienced \emph{either} proximal or personal contact \emph{at any level of intensity}, including being stopped and questioned by police, were mobilized to take part in nonvoting political activity---and the same held true for whites, but with a much smaller effect size. Walker theorizes that this finding suggests that nonwhite people view criminalization ``as a form of systemic political threat,'' an analysis comparable to ``linked fate.'' Notably, Walker's survey questions excluded ``minor traffic stops,'' the primary focus of this paper.

Ariel White leveraged administrative court records from Harris County, Texas, to identify voters who lived within individuals who were charged with a misdemeanor offense and arrested, convicted, or jailed (\protect\hyperlink{ref-White2019}{White 2019a}). She found that living with someone who was charged but not convicted in the 30 day period preceding an election increased voter turnout by 3 percentage points, whereas ``convicted but not jailed'' decreased turnout by 6 percentage points and ``convicted and jailed'' decreased turnout by 12 percentage points. This provides further support for the theory that increasing intensity of criminalization could change the direction of the effect of criminalization on political behavior---and, in fact, that lower-level contact can mobilize even as higher-level contact demobilizes.

Although geographic or spatial analyses of how criminalization affects political socialization technically move beyond proximal contact as an extant conceptual framework, we argue that geographic proximity should work similarly to social network proximity---if one neighborhood is heavily policed, people who live in that neighborhood are more likely to have social ties who have experienced policing. Traci Burch found that North Carolina block groups with higher concentrations of people who had experienced incarceration or criminal legal supervision saw substantial reductions in voter turnout, and confirmed her findings at the individual level (\protect\hyperlink{ref-Burch2014}{Burch 2014}). \protect\hyperlink{ref-Morris2020}{Morris} (\protect\hyperlink{ref-Morris2020}{2020}) similarly finds that neighborhoods with more active voters lost to felony disenfranchisement turned out at lower rates in the 2017 mayoral race in NYC, and that this effect was particularly pronounced in Black neighborhoods.

@@

Over the past few years, scientific understanding of the political effects of more serious interactions with the criminal legal system has grown by leaps and bounds. Table \ref{tab:lit-table} summarizes our understanding of the literature. As Table \ref{tab:lit-table} makes clear, a large body of research has now demonstrated the negative effects of high-level contact with the criminal legal system on voter turnout. Although lower-level stops are far more frequent, our knowledge of their (de)mobilizing effects is under-developed.

\begin{singlespace}
\input{../raw_data/lit_tab.tex}
\end{singlespace}

Most of the aforementioned studies were conducted using data collected before the rise of the Black Lives Matter movement in 2013, which could be construed as a significant development for the political socialization of Black Americans with respect to policing and incarceration. The movement garnered national attention and momentum during the Ferguson protests in 2014, where racially discriminatory ticketing practices were particularly egregious. Though racist police violence was and is a focal point for many Black Lives Matter actions, local chapters frequently use policing as an entry point into a wider set of demands regarding the material security of Black Americans and their marginalization in American politics---and these demands affect police practices. One recent study found that census places that saw Black Lives Matter protests experienced a 15 to 20 percent decrease in police killings over the following five year period (\protect\hyperlink{ref-Campbell2021}{Campbell 2021}). It's possible that any potentially mobilizing effects of criminal legal contact became stronger in the 2010---in other words, Black Lives Matter could have facilitated a more robust and prevalent linked-fate political analysis among Black Americans during the time frame of our analysis. That said, \protect\hyperlink{ref-Laniyonu2019}{Laniyonu} (\protect\hyperlink{ref-Laniyonu2019}{2019}) did demonstrate mobilization in Black neighborhoods before the beginning of the Black Lives Matter movement.

@@

In the 20th century, residential segregation simultaneously fueled the racial wealth gap and ensured that criminalization was both spatially and racially concentrated (\protect\hyperlink{ref-Rothstein2017}{Rothstein 2017}; \protect\hyperlink{ref-Hinton2021}{Hinton and Cook 2021}). At the level of federal policy, racial threat and whites' attribution of criminality to Blackness (\protect\hyperlink{ref-Muhammad2010}{Muhammad 2010}) prompted a shift ``from the War on Poverty to the War on Crime'' wherein Black communities' demands for material security and safety were selectively amplified to promote increased and harsher policing and incarceration (\protect\hyperlink{ref-Hinton2017}{Hinton 2017}; \protect\hyperlink{ref-Vargas2017a}{Vargas and McHarris 2017}). Ruth Wilson Gilmore explains how mid to late 20th-century whites ``demanded that the state become leaner and meaner, except when directed to do otherwise''---in other words, Keynesian welfare guarantees were undermined by the racist myth of ``welfare queens'' and supplanted by expanded police power (\protect\hyperlink{ref-Gilmore2007}{Gilmore 2007, 53}).

The 2008 financial crisis solidified the connection between fiscal austerity and policing, prompting a decline in tax revenues that devastated state and local governments---but law enforcement agencies were largely spared from cuts. At the local level, this means that police budgets today constitute about thirty percent of city budgets nationwide, and cities where residents have lower median incomes dedicate a larger proportion of spending to policing (\protect\hyperlink{ref-Friedman2020}{Friedman 2020}). In municipalities with substantial Black populations, then, people who are most heavily affected by criminalization are expected to foot the bill for their own mistreatment.

@@

In 2020, a Federal Reserve survey found that five percent of white respondents reported ``unpaid legal expenses, fines, or court costs,'' whereas twelve percent of Black respondents reported such debt (\protect\hyperlink{ref-BoardofGovernorsoftheFederalReserveSystem2020}{Governors of the Federal Reserve System 2020}). Across the country, city government reliance on fines and fees revenue is associated with larger Black populations (\protect\hyperlink{ref-UnitedStatesCommissiononCivilRights2017}{United States Commission on Civil Rights 2017}). Although major urban centers collect high dollar amounts of fines and fees, some of the most egregious ticketing practices measured on a per-capita basis are concentrated in suburbs or smaller cities with municipal courts---Ferguson being a prominent example (\protect\hyperlink{ref-UnitedStatesDepartmentofJusticeCivilRightsDivision2015}{United States Department of Justice Civil Rights Division 2015}; \protect\hyperlink{ref-Mughan2020}{Mughan 2020}). The positive association between cities' reliance on fines and fees and the proportion of Black residents is significantly diminished when Black communities are represented by at least one Black legislator, however (\protect\hyperlink{ref-Sances2017}{Sances and You 2017}).

@@

Research regarding police stops and searches in response to local fiscal crisis shows how decreasing local revenues affect racist police practices. One study of all U.S. city governments found that local fiscal stress was related to increased ``revenue-generating arrests'' (e.g.~arrests involving civil asset forfeiture) and that Black and Hispanic people were arrested at higher rates (\protect\hyperlink{ref-Makowsky2009}{Makowsky and Stratmann 2009}). A study of four million North Carolina traffic stops found that fiscal stress caused a decrease in the probability of white drivers being searched, but no decrease for Black drivers (\protect\hyperlink{ref-Shoub2020}{Shoub et al. 2020}). This suggests that higher reliance on fines and fees results in government officials and police targeting individuals who hold less political power---whether those individuals are out-of-state motorists or members of local marginalized communities (\protect\hyperlink{ref-Makowsky2009}{Makowsky and Stratmann 2009}).

@@

We merge these geocoded voter files with ``place'' and ``county subdivision'' shapefiles published by the US Census Bureau. By aggregating these individual-level records to the municipal level, we are able to estimate the number of registered voters and the number of ballots cast in every city in the country. By using the racial estimates provided by L2, we can estimate the number of ballots cast by voters of each race in each city. Although each voters' race is modeled based on L2's algorithms---and is not self-reported---it is possibly incorrect for some voters. We assume this does not present a problem when records are aggregated to the municipal level.

@@

\emph{Hypothesis 2: The effect of the city-level relationship between fines and voter turnout is less demobilizing or more mobilizing for Black voters, given the historical and explicit racial injustice narratives used to explain these practices.}

@@

To test the effect of ticketing on individual-level turnout, we match the first and last names and dates of birth against the Hillsborough County registered voter file. Here, rather than rely on a snapshot from L2, we use the raw voter file directly from the state. Using a single post-treatment snapshot of the voter file can result in conditioning on a post-treatment status (see \protect\hyperlink{ref-Nyhan2017}{Nyhan, Skovron, and Titiunik 2017}). Instead, we collect snapshots of the voter file following each even-year general election between 2012 and 2018. We thus observe virtually all individuals who were registered to vote at any time during our period of study. Unique voter identification numbers allow us to avoid double-counting voters who are registered in multiple snapshots. We retain each voter's earliest record, and geocode voters to their home census block groups.

@@

The use of administrative data marks an important step forward in our understanding of how low-level contact with the criminal legal system structures participation. Past work looking at the individual-level effects of low-level contact has relied entirely on survey or qualitative data (e.g. \protect\hyperlink{ref-Walker2014}{Walker 2014}; \protect\hyperlink{ref-Weaver2010}{Weaver and Lerman 2010}). Their in-depth research allows for the testing of specific psychological mechanisms and personal interpretations of criminal legal contact, but does not allow us to generalize more broadly. As Weaver and Lerman (\protect\hyperlink{ref-Weaver2010}{2010, 821}) note, it may also introduce measurement bias. Our analysis investigates actual voting behavior following actual ticketing events, not reported voting behavior or reported exposure to police ticketing. The administrative data therefore allow us both to sidestep reporting error and to observe the behavior of the 232,568 individuals ticketed over the 6 year period---a far larger pool than even the most robust surveys.

@@

Figure \ref{fig:did2} makes this effect even clearer. Here, we limit the analysis to voters who were stopped within the 90 days prior to an election (and their controls). We expected that the ticket would be more salient for individuals stopped shortly before the election. This appears to be the case---\emph{especially for Black voters.}

@@

\begin{singlespace}
\input{"../temp/dind_reg_y.tex"}
\end{singlespace}
